### 2017-10-17 SNEQ_Lynott_example.Rmd
Completed first draft of the significant non-equivalent proportions test from Lynott et al. Skipped the reference to values from Williams & Barge, since they are not used in any calcuations, and rephrased the bound criterion to smallest effect size that we can detect given the current sample. An alternative would be to calculate the smallest detectable effect size given the original sample.

---
title: "SNEQ_Lynott_example"
author: "Peder M Isager"
date: "October 17, 2017"
output: html_document
---

# Example of effect that is significant and not equivalent

Lynott et al. (2014), replicating Williams & Bargh (2008) conducted a study to investigate the effect of being exposed to physical warmth on interpersonal warmth and prosocial behavior. They observed  that 51.2% of participants who received a cold pack (n = 427) opted to receive a reward for themselves, while 57.9% (n = 434) of participants who received a warm pack did the same. Calculating a z-test for the difference between the proportions, this effect is statistically significant (Diff = -0.067, z = -1.98, p = 0.024). Having obtained a significant result, we are left with two alternatives:

1. The difference is small enough that even though it is *statisitcally* significant, we can consider it *practically* equivalent to zero.
2. The difference is substantial enough that we cannot rule out that it is also practically significant.

In order to distinguish between these two hypotheses, we calculate the TOST. We first need to define our SESOI (and thus, our bounds). In this case, it is not really clear what an objective cirterion for the smallest effect size of interest could be. However, we could decide that we do not care about an effect if the difference is smaller than the smallest effect we could potentially detect with the current sample. This essentially means that we use the critical z value (~1.96 in a two-tailed test with an alpha of 0.05) as our bounds. To convert the critical z value to the critical difference in proportions for this study, we multiply critical z with the standard error:

```{r}
se     <- sqrt((0.512 * (1 - 0.512)) / 427 + (0.579 * (1 - 0.579)) / 434)  # Calculate the (unpooled) standard error
crit.z <- qnorm(1 - (0.05 / 2))  # Determine the critical z for a two-tailed test with alpha = 0.05

crit.diff <- crit.z * se  # Convert z to the critical difference

print(crit.diff)
```

Using the TOSTtwo.prop functiion in the TOSTER package in R, we can then calculate the TOST by insterting the summary statistics from Lynott et al. (2014), assuming an aplha of 0.05, and using +/- the value of `crit.diff` (`r crit.diff`) as our equivalence bounds. We find that the TOST is non-significant, t(96.62) = 0.72, p = 0.24. Thus, in addition to begin statistically different from zero, we cannot reject the possibility that the true difference really is larger than our SESOI.

```{r}

download.file(url = "https://raw.githubusercontent.com/Lakens/EquivalenceTestingForPsychologicalScience/master/TOSTtwo.prop.R", destfile = "TOSTtwo.prop.R")
source(file = "TOSTtwo.prop.R")

TOSTtwo.prop(prop1 = 0.512, 
             prop2 = 0.579, 
             n1    = 427, 
             n2    = 434, 
             low_eqbound  = -crit.diff, 
             high_eqbound = crit.diff, 
             alpha = 0.05)

```