### 2017-10-18 NSNEQ_Moon_Example.Rmd
My last edits to Moon example before Daniel's comments. Remoced details of original as it was unnecessary for the example.

---
title: "NSNEQ_Moon_Example"
author: "Peder M Isager"
date: "2017 M10 16"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Example of effect that is neither significant nor equivalent

Moon & Roeder (2014), replicating Shih, Pittinsky, and Ambady (1999), conducted a study to investigate whether Asian-American women would perform better on a maths test when primed with their Asian identity. They observed a slightly reversed difference between the Asian primed group (N = 53, M = 0.46, SD = 0.17) and the control (N = 48, M = 0.50, SD - 0.18) which was not significant, d = -0.23, Welch t(96.62) = 1.15, p = 0.26, two-sided). Having failed to obtain a significant result, we are left with two alternatives:

1. The data indicate the absence of any effect, and we can accept the null hypothesis.
2. The data are not sensitive enough to tell us whether an effect is present or abscent.

In order to distinguish between these two hypotheses, we calculate the TOST. We first need to define our SESOI (and thus, our bounds). Let us imagine that we consult a teacher about the test, and discover that that grades for this test are set at every 6.25% increase in correct answers (F = 0% to 6.25% ... A+ = 93.75% to 100%). We can decide that we are only interested in test differences so far as they correspond to at least a 1 gradepoint increase or decrease. We can convert this percentage difference into a corresponding Cohen's d for Moon & Roeder's (2014) data by multiplying 6.25/100 on the pooled standard deviation of the groups.

$$d = \frac{Diff}{SD_{pooled}}$$  
$$SD_{pooled} = \sqrt \frac{sd_{1}^2 + sd_{2}^2}{2}$$

``` {r}

sesoi     <- 0.065                        # <- Set the smallest effect size of interest
sd.pooled <- sqrt((0.17^2 + 0.18^2) / 2)  # <- Calculate the pooled standard deviation
sesoi.d   <- 0.065 / sd.pooled            # <- Convert the raw SESOI to Cohen's d

print(sesoi.d)

```

Using the TOSTtwo function in the TOSTER package in R, we can then calculate the TOST by insterting the summary statistics from Moon & Roeder (2014), assuming an aplha of 0.05, and using +/- the value of `sesoi.d` (`r sesoi.d`) as our equivalence bounds. We find that the TOST is non-significant, t(96.62) = 0.72, p = 0.24. Thus, we cannot reject the possibility that the true effect really is larger than our SESOI.

```{r}

# Below are the code with rounded and ~accurate values. We need to decide which to use. The accurate one will correspond better to Jamovi. The rounded one will correspont better to the values given above. I (Peder) prefer the rounded.
TOSTER::TOSTtwo(m1  = 0.46,
                m2  = 0.50,
                sd1 = 0.17,
                sd2 = 0.18,
                n1  = 53,
                n2  = 48,
                low_eqbound  = -sesoi.d,
                high_eqbound = sesoi.d,
                alpha        = .05)

TOSTER::TOSTtwo(m1  = 0.4585130,
                m2  = 0.4962693,
                sd1 = 0.1749011,
                sd2 = 0.1770452,
                n1  = 53,
                n2  = 48,
                low_eqbound  = -sesoi.d,
                high_eqbound = sesoi.d,
                alpha        = .05)
```

```{r}
#Jamovi results - This is just included to illustrate the correspondence between R results and what readers will see in Jamovi

Moon_data <- read.csv("Moon_data_asianvscontrol.csv")
Moon_data_edit <- Moon_data[!Moon_data$identity_salience==3,]

TOSTER::dataTOSTtwo(data  = Moon_data_edit,
                    deps  = "accuracy",
                    group = "identity_salience",
                    low_eqbound_d  = -sesoi.d,
                    high_eqbound_d = sesoi.d,
                    plots     = TRUE,
                    var_equal = FALSE)
```