### 2017-10-18 NSNEQ_Moon_Example.Rmd

---
title: "NSNEQ_Moon_Example"
author: "Peder M Isager"
date: "2017 M10 16"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
<!-- I don't think we need to say 'Welch's t-test. We might throw in a quick remakr that by default, the TOST package reports Welch's t-test (following Delacre et al., 2017) -->  
# Example of an effect that is neither statistically significant nor equivalent

Moon & Roeder (2014), replicating Shih, Pittinsky, and Ambady (1999), conducted a study to investigate whether Asian-American women would perform better on a maths test when primed with their Asian identity. They observed a slightly reversed difference between the Asian primed group (N = 53, M = 0.46, SD = 0.17) and the control (N = 48, M = 0.50, SD - 0.18) which was not significant, d = -0.23, Welch t(96.62) = 1.15, p = 0.26, two-sided). Having failed to obtain a significant result, we are left with two alternatives:

1. The data indicate the absence of a meaningful effect, and we can accept the null hypothesis.
2. The data are not sensitive enough to tell us whether a meaningful effect is present or absent.

In order to distinguish between these two hypotheses, we first need to define a SESOI (and thus, our equivalence bounds). Let us imagine that we consult a teacher about the test, and discover that that grades for this test are set at every 6.25% increase in correct answers (F = 0% to 6.25% ... A+ = 93.75% to 100%). We can decide that we are only interested in test differences so far as they correspond to at least a 1 gradepoint increase or decrease. We can convert this percentage difference into a corresponding Cohen's d for Moon & Roeder's (2014) data by multiplying 6.25/100 on the pooled standard deviation of the groups.

$$d = \frac{Diff}{SD_{pooled}}$$  
$$SD_{pooled} = \sqrt \frac{sd_{1}^2 + sd_{2}^2}{2}$$

``` {r}

sesoi     <- 0.065                        # <- Set the smallest effect size of interest
sd.pooled <- sqrt((0.17^2 + 0.18^2) / 2)  # <- Calculate the pooled standard deviation
sesoi.d   <- 0.065 / sd.pooled            # <- Convert the raw SESOI to Cohen's d

print(sesoi.d)

```  
<!-- So the TOST results below should not be typed in, but calculated in R, like you do with sesoi.d -->

Using the TOSTtwo function in the TOSTER package in R, we can then calculate the TOST by insterting the summary statistics from Moon & Roeder (2014), assuming an aplha of 0.05, and using +/- the value of `sesoi.d` (`r sesoi.d`) as our equivalence bounds. We find that the TOST is non-significant, t(96.62) = 0.72, p = 0.24. Thus, we cannot reject the possibility that the true effect really is larger than our SESOI.

```{r}

# Below are the code with rounded and ~accurate values. We need to decide which to use. The accurate one will correspond better to Jamovi. The rounded one will correspont better to the values given above. I (Peder) prefer the rounded.
TOSTER::TOSTtwo(m1  = 0.46,
                m2  = 0.50,
                sd1 = 0.17,
                sd2 = 0.18,
                n1  = 53,
                n2  = 48,
                low_eqbound  = -sesoi.d,
                high_eqbound = sesoi.d,
                alpha        = .05)

TOSTER::TOSTtwo(m1  = 0.4585130,
                m2  = 0.4962693,
                sd1 = 0.1749011,
                sd2 = 0.1770452,
                n1  = 53,
                n2  = 48,
                low_eqbound  = -sesoi.d,
                high_eqbound = sesoi.d,
                alpha        = .05)
```

```{r}
#Jamovi results - This is just included to illustrate the correspondence between R results and what readers will see in Jamovi

Moon_data <- read.csv("Moon_data_asianvscontrol.csv")
Moon_data_edit <- Moon_data[!Moon_data$identity_salience==3,]

TOSTER::dataTOSTtwo(data  = Moon_data_edit,
                    deps  = "accuracy",
                    group = "identity_salience",
                    low_eqbound_d  = -sesoi.d,
                    high_eqbound_d = sesoi.d,
                    plots     = TRUE,
                    var_equal = FALSE)
```