### 2017-10-30 NSNEQ_Moon_Example.Rmd
Added footnote
* Added footnote conserning the difference between using a raw effect, the Cohen's d, and the Cohen's d based on the SD of another study.

---
title: "NSNEQ_Moon_Example"
author: "Peder M Isager"
date: "2017 M10 16"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Example of an effect that is neither statistically significant nor equivalent

Moon & Roeder (2014), replicating Shih, Pittinsky, and Ambady (1999), conducted a study to investigate whether Asian-American women would perform better on a maths test when primed with their Asian identity. They observed a slightly reversed difference between the Asian primed group (N = 53, M = 0.46, SD = 0.17) and the control (N = 48, M = 0.50, SD - 0.18) which was not significant, d = -0.23, t(96.62) = 1.15, p = 0.26, two-sided). Having failed to obtain a significant result, we are left with two alternatives:

1. The data indicate the absence of a meaningful effect, and we can accept the null hypothesis.
2. The data are not sensitive enough to tell us whether a meaningful effect is present or absent.

In order to distinguish between these two hypotheses, we first need to define a SESOI (and thus, our equivalence bounds). Let us imagine that we consult a teacher about the test, and discover that that grades for this test are set at every 6.25% increase in correct answers (F = 0% to 6.25% ... A+ = 93.75% to 100%). We can decide that we are only interested in test differences so far as they correspond to at least a 1 gradepoint increase or decrease. Thus, our SESOI becomes 6.25%, or 0.0625.

Using the TOSTtwo function in the TOSTER package in R, we can then calculate the TOST by insterting the summary statistics from Moon & Roeder's (2014) data, assuming an aplha of 0.05, and using +/- the SESOI of 0.0625 as our equivalence bounds.

```{r echo=FALSE}

#  We first load the relevant datafile into R
moon.data    <- read.csv("NSNEQ_Moon_Example_data.csv")

#  We then calculate the summary statistics from the data that we need in the TOST calculation below
mean.asian   <- mean(moon.data$accuracy[moon.data$identity_salience == 1])
mean.control <- mean(moon.data$accuracy[moon.data$identity_salience == 2])
sd.asian     <- sd(moon.data$accuracy[moon.data$identity_salience == 1])
sd.control   <- sd(moon.data$accuracy[moon.data$identity_salience == 2])
n.asian      <- sum(moon.data$identity_salience == 1)
n.control    <- sum(moon.data$identity_salience == 2)

#  We can now run the TOST!
#  Performing the TOST using the 'TOSTtwo.raw' function allows us to specify equivalence bounds in raw values
tost <- TOSTER::TOSTtwo.raw(m1  = mean.asian,
                            m2  = mean.control,
                            sd1 = sd.asian,
                            sd2 = sd.control,
                            n1  = n.asian,
                            n2  = n.control,
                            low_eqbound  = -0.0625,
                            high_eqbound =  0.0625,
                            alpha = .05)
```

We find that the TOST is non-significant, t(`r round(tost$TOST_df, 2)`) = `r round(min(abs(c(tost$TOST_t1, tost$TOST_t2))), 2)`, p = `r round(max(tost$TOST_p1, tost$TOST_p2), 2)`. Thus, we cannot reject the possibility that the true effect really is larger than our SESOI.[^footnote]

[^footnote]: In this example, the bounds of the TOST was set based on a raw effect of 6.25%. We could have converted this raw effect into a standardized effect (Cohen's d) for the current sample, and we would have gotten the exact same result, only then in Cohen's d (equivalence bounds = +/- `r round(0.0625 / sqrt((sd.asian^2 + sd.control^2) / 2), 2)`). Since this is a replication of a previous study, we could also have chosen calculate the Cohen's d that corresponds to the raw SESOI in the *original* study, and used this as the bounds for the replication. Now we would no longer be answering a question about a raw effect independent of the samples. Instead, we are asking whether we in the replication can reject standardized effects larger than a certain standardized effect in the original study. Since the standard deviations of the original study were only slightly higher than in the replication in this case, Cohen's d, and thus the bounds, would also be only sligthly smaller (bounds = +/- `r round(0.0625 / sqrt((0.17^2 + 0.20^2) / 2), 2)`). As standard deviations of the original study compared to the replication becomes larger, the bounds become more narrow.